Illusion Sensitivity and Phenomenological Control

### 2025-05-21 analysis/2_analysis.qmd
lesgooo @AnafNeves

---
title: "Illusion Sensitivity and Phenomenological Control"
subtitle: "Analysis"
date: "`r Sys.Date()`"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
    warning: false 
    message: false 
execute: 
  cache: true
---

@@

data_ig <- read.csv("../data/data_IllusionGame.csv") |>
  group_by(Illusion_Type) |> 
  mutate(Illusion_Side = sign(Illusion_Difference), 
    Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
    Task_Difficulty = abs(Illusion_Difference),
      Illusion_Group = case_when(
        Illusion_Effect == "Congruent" ~ "Congruent",
        Illusion_Effect == "Incongruent" & Illusion_Strength <= median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Mild",
        Illusion_Effect == "Incongruent" & Illusion_Strength > median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Strong"),
    Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
  ungroup() |>
  mutate(
    Block = as.factor(Block),
    Illusion_Side = as.factor(Illusion_Side),
    Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
    Response_Correct = as.factor(Response_Correct)
  )

@@

## Illusion Sensitivity Scores

### Error Rate

```{r}
errorrate <- data_ig |>
  group_by(Participant, Illusion_Type, Illusion_Group) |>
  summarize(ErrorRate = sum(Error) / n()) |>
  ungroup() |>
  as.data.frame()
```

### IES

**Note.** The IES is calculated by dividing the *average correct* RT by the *proportion of correct responses*, for each illusion and for each illusion strenght group: Strong Illusion Strength & Incongruent, Mild Illusion Strength & Incongruent, and Congruent

```{r}

IES <- data_ig |>
  group_by(Participant, Illusion_Type, Illusion_Group) |>
  summarise(IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true"))) |>
  ungroup() |>
  as.data.frame()
```

```{r}

data_vs <- left_join(IES, errorrate, by= c("Participant", "Illusion_Type", "Illusion_Group"))
```

### Correlations between Strong and Mild

**Note**.  If correlations high between mild and strong conditions then the mild and strong illusion strength groups will be collapsed and the outcomes will be recomputed across all trials, otherwise they will be treated as separate in subsequent analyses.

```{r}

compute_correlation <- function(data, illusion_type) {
  illusion_data <- data |> dplyr::filter(Illusion_Group != "Congruent", Illusion_Type == illusion_type) |>
    pivot_wider(id_cols = Participant,
      names_from = Illusion_Group,
      values_from = c(ErrorRate, IES),
      names_sep = "_")

  ER_Incongruent <- BayesFactor::correlationBF(
    y = illusion_data$ErrorRate_Mild, 
    x = illusion_data$ErrorRate_Strong
  )

  IES_Incongruent <- BayesFactor::correlationBF(
    y = illusion_data$IES_Mild, 
    x = illusion_data$IES_Strong
  )

  return(list(
    ER_Incongruent = ER_Incongruent,
    IES_Incongruent = IES_Incongruent
  ))
}

```{r}

cor_ebbinghaus <- compute_correlation(data_vs, "Ebbinghaus")
cor_mullerlyer <- compute_correlation(data_vs, "MullerLyer")
cor_verticalhorizontal <- compute_correlation(data_vs, "VerticalHorizontal")
```

```{r}

er_ebbinghaus_post<- describe_posterior(cor_ebbinghaus$ER_Incongruent)
er_mullerlyer_post <- describe_posterior(cor_mullerlyer$ER_Incongruent)
er_verticalhorizontal_post <- describe_posterior(cor_verticalhorizontal$ER_Incongruent)
```

```{r}

ies_ebbinghaus_post <- describe_posterior(cor_ebbinghaus$IES_Incongruent)
ies_mullerlyer_post <- describe_posterior(cor_mullerlyer$IES_Incongruent)
ies_verticalhorizontal_post <- describe_posterior(cor_verticalhorizontal$IES_Incongruent)

### Colapsed scores

```{r}

# collapsed_groups_data <- data_vs |>
#  filter(Illusion_Group %in% c("Mild", "Strong")) |>
#   group_by(Participant, Illusion_Type) |>
#   summarise(
#     Illusion_Group = "Incongruent",
#     ErrorRate = mean(ErrorRate, na.rm = TRUE),
#     IES = mean(IES, na.rm = TRUE),
#     .groups = "drop"
#   )

# data_vs <- bind_rows(data_vs, collapsed_groups_data) |>
#   arrange(Participant, Illusion_Type, factor(Illusion_Group, levels = c("Mild", "Strong", "Congruent", "Incongruent"))) |>
#   filter(Illusion_Group == c( "Congruent", "Incongruent"))

```{r}
data_vs_wide <- data_vs|>
  select(Participant, Illusion_Type, Illusion_Group, ErrorRate, IES) |>
  unite("group", Illusion_Type, Illusion_Group, sep = "_") |>
  pivot_wider(names_from = group, values_from = c(ErrorRate, IES)) |>
  as.data.frame()
```

### Illusion Manipulation Checks

**Note:** Significant differences between the Congruent and Mild and the Mild and Strong groups will be taken as evidence that illsuions functioned as intended.

```{r}
compute_ttest <- function(data, illusion_type) {
  illusion_data <- data |> dplyr::filter(Illusion_Type == illusion_type)

  ERcongruent_vs_mild <- BayesFactor::ttestBF(
    x = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Congruent"],
    y = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Mild"],
    paired = TRUE
  )

  IEScongruent_vs_mild <- BayesFactor::ttestBF(
    x = illusion_data$IES[illusion_data$Illusion_Group == "Congruent"],
    y = illusion_data$IES[illusion_data$Illusion_Group == "Mild"],
    paired = TRUE
  )

  ERmild_vs_strong <- BayesFactor::ttestBF(
    x = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Mild"],
    y = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Strong"],
    paired = TRUE
  )

  IESmild_vs_strong <- BayesFactor::ttestBF(
    x = illusion_data$IES[illusion_data$Illusion_Group == "Mild"],
    y = illusion_data$IES[illusion_data$Illusion_Group == "Strong"],
    paired = TRUE
  )

  # Congruent vs Incongruent if correlation high between mild and strong

  # ERcong_vs_incon < BayesFactor::ttestBF(
  #   x = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Congruent"],
  #   y = illusion_data$ErrorRate[illusion_data$Illusion_Group == "Incongruent"],
  #   paired = TRUE
  # )
  # 
  # IEScong_vs_incon <- BayesFactor::ttestBF(
  #   x = illusion_data$IES[illusion_data$Illusion_Group == "Congruent"],
  #   y = illusion_data$IES[illusion_data$Illusion_Group == "Incongruent"],
  #   paired = TRUE
  # )

  return(list(
    ERcongruent_vs_mild = ERcongruent_vs_mild,
    ERmild_vs_strong = ERmild_vs_strong,
    IEScongruent_vs_mild = IEScongruent_vs_mild,
    IESmild_vs_strong = IESmild_vs_strong
    # ERcong_vs_incon = ERcong_vs_incon,
    # IEScong_vs_incon = IEScong_vs_incon
  ))
}

```{r}

ttest_ebbinghaus <- compute_ttest(data_vs, "Ebbinghaus")
ttest_mullerlyer <- compute_ttest(data_vs, "MullerLyer")
ttest_verticalhorizontal <- compute_ttest(data_vs, "VerticalHorizontal")
```

```{r}

er_ebbinghaus_post_cm <- describe_posterior(ttest_ebbinghaus$ERcongruent_vs_mild)
er_ebbinghaus_post_ms <- describe_posterior(ttest_ebbinghaus$ERmild_vs_strong)
# er_ebbinghaus_post_ci <- describe_posterior(ttest_ebbinghaus$ERcong_vs_incon)

er_mullerlyer_post_cm <- describe_posterior(ttest_mullerlyer$ERcongruent_vs_mild)
er_mullerlyer_post_ms <- describe_posterior(ttest_mullerlyer$ERmild_vs_strong)
# er_mullerlyer_post_ci <- describe_posterior(ttest_mullerlyer$ERcong_vs_incon)

er_verticalhorizontal_post_cm <- describe_posterior(ttest_verticalhorizontal$ERcongruent_vs_mild)
er_verticalhorizontal_post_ms <- describe_posterior(ttest_verticalhorizontal$ERmild_vs_strong)
# er_verticalhorizontal_post_ci <- describe_posterior(ttestcor_verticalhorizontal$ERcong_vs_incon)

```{r}

IES_ebbinghaus_post_cm <- describe_posterior(ttest_ebbinghaus$IEScongruent_vs_mild)
IES_ebbinghaus_post_ms <- describe_posterior(ttest_ebbinghaus$IESmild_vs_strong)
# IES_ebbinghaus_post_ci <- describe_posterior(ttest_ebbinghaus$IEScong_vs_incon)

IES_mullerlyer_post_cm <- describe_posterior(ttest_mullerlyer$IEScongruent_vs_mild)
IES_mullerlyer_post_ms <- describe_posterior(ttest_mullerlyer$IESmild_vs_strong)
# IES_mullerlyer_post_ci <- describe_posterior(ttest_ebbinghaus$IEScong_vs_incon)

IES_verticalhorizontal_post_cm <- describe_posterior(ttest_verticalhorizontal$IEScongruent_vs_mild)
IES_verticalhorizontal_post_ms <- describe_posterior(ttest_verticalhorizontal$IESmild_vs_strong)
# IES_verticalhorizontal_post_ci <- describe_posterior(ttest_verticalhorizontal$IEScong_vs_incon)
```

### Reliability

**Note**. Reliability analyses are conducted on all resulting indices.

#### Split-half

```{r}

compute_all_split_half <- function(df, score_vars = c("ErrorRate", "IES")) {

  conditions <- c("Congruent", "Strong", "Mild")
  # conditions <- c("Congruent", "Incongruent")
  illusions <- c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")
  pairs <- list(
    c("Ebbinghaus", "MullerLyer"),
    c("VerticalHorizontal", "MullerLyer"),
    c("VerticalHorizontal", "Ebbinghaus")
  )

  # Loop through score vars and compute
  results <- list()

  for (score in score_vars) {
    cat("\n###", score, "\n")

    for (cond in conditions) {
      cat("\n##", cond, "\n")

      for (pair in pairs) {
        var1 <- paste(score, pair[1], cond, sep = "_")
        var2 <- paste(score, pair[2], cond, sep = "_")

        if (all(c(var1, var2) %in% colnames(df))) {
          cat("Pair:", var1, "vs", var2, "\n")
          res <- performance::item_split_half(df[, c(var1, var2)])
          results[[paste(score, cond, paste(pair, collapse = "_"), sep = "_")]] <- res
        } else {
          warning(paste("Missing columns:", var1, "or", var2))
        }
      }
    }
  }

  return()
}

split_half_results  <- compute_all_split_half(data_vs_wide)

#### Inter-illusion Reliability

Error Rate

```{r}

con_rel_er <- data_vs_wide |>
  select(matches("ErrorRate.*_Congruent")) |>
  psych::alpha()

mild_rel_er <- data_vs_wide |>
  select(matches("ErrorRate.*_mild")) |>
  psych::alpha()

strong_rel_er <- data_vs_wide |>
  select(matches("ErrorRate.*_strong")) |>
  psych::alpha()

# inc_rel_er <- data_vs_wide |>
#   select(matches("ErrorRate.*_Inongruent")) |>
#   psych::alpha()

```{r}

con_rel_IES <- data_vs_wide |>
  select(matches("IES.*_Congruent")) |>
  psych::alpha()

mild_rel_IES <- data_vs_wide |>
  select(matches("IES.*_mild")) |>
  psych::alpha()

strong_rel_IES <- data_vs_wide |>
  select(matches("IES.*_strong")) |>
  psych::alpha()

# inc_rel_IES <- data_vs_wide |>
#   select(matches("IES.*_Inongruent")) |>
#   psych::alpha()

## PCS and Illusion Sensitivity

### Error Rate

#### Ebbinghaus 
```{r}

cor_pcs_er_ebb_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_Ebbinghaus_Congruent)

er_pcs_ebb_con_post <- describe_posterior(cor_pcs_er_ebb_con)

cor_pcs_er_ebb_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_Ebbinghaus_Mild)

er_pcs_ebb_mild_post <- describe_posterior(cor_pcs_er_ebb_m)

cor_pcs_er_ebb_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_Ebbinghaus_Strong)

er_pcs_ebb_str_post <- describe_posterior(cor_pcs_er_ebb_s)

# Incongruent
# cor_pcs_er_ebb_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_Ebbinghaus_Incongruent)
# 
# er_pcs_ebb_inc_post <- describe_posterior(cor_pcs_er_ebb_inc)

#### Muller-Lyer
```{r}

cor_pcs_er_mul_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_MullerLyer_Congruent)

er_pcs_mul_con_post <- describe_posterior(cor_pcs_er_mul_con)

cor_pcs_er_mul_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_MullerLyer_Mild)

er_pcs_mul_mild_post <- describe_posterior(cor_pcs_er_mul_m)

cor_pcs_er_mul_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_MullerLyer_Strong)

er_pcs_mul_str_post <- describe_posterior(cor_pcs_er_mul_s)

# Incongruent
# cor_pcs_er_mul_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_MullerLyer_Incongruent)
# 
# er_pcs_mul_inc_post <- describe_posterior(cor_pcs_er_mul_inc)

#### VerticalHorizontal
```{r}

cor_pcs_er_vh_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_VerticalHorizontal_Congruent)

er_pcs_vh_con_post <- describe_posterior(cor_pcs_er_vh_con)

cor_pcs_er_vh_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_VerticalHorizontal_Mild)

er_pcs_vh_mild_post <- describe_posterior(cor_pcs_er_vh_m)

cor_pcs_er_vh_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$ErrorRate_VerticalHorizontal_Strong)

er_pcs_vh_str_post <- describe_posterior(cor_pcs_er_vh_s)

# Incongruent
# cor_pcs_er_vh_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_VerticalHorizontal_Incongruent)
# 
# er_pcs_vh_inc_post <- describe_posterior(cor_pcs_er_vh_inc)
```

### IES

#### Ebbinghaus 
```{r}

cor_pcs_IES_ebb_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$IES_Ebbinghaus_Congruent)

IES_pcs_ebb_con_post <- describe_posterior(cor_pcs_IES_ebb_con)

cor_pcs_IES_ebb_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$IES_Ebbinghaus_Mild)

IES_pcs_ebb_mild_post <- describe_posterior(cor_pcs_IES_ebb_m)

cor_pcs_IES_ebb_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs,
  x = data_vs_wide$IES_Ebbinghaus_Strong)

IES_pcs_ebb_str_post <- describe_posterior(cor_pcs_IES_ebb_s)

# Incongruent
# cor_pcs_IES_ebb_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_Ebbinghaus_Incongruent)
# 
# IES_pcs_ebb_inc_post <- describe_posterior(cor_pcs_IES_ebb_inc)

#### Muller-Lyer
```{r}

cor_pcs_IES_mul_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_MullerLyer_Congruent)

IES_pcs_mul_con_post <- describe_posterior(cor_pcs_IES_mul_con)

cor_pcs_IES_mul_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_MullerLyer_Mild)

IES_pcs_mul_mild_post <- describe_posterior(cor_pcs_IES_mul_m)

cor_pcs_IES_mul_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_MullerLyer_Strong)

IES_pcs_mul_str_post <- describe_posterior(cor_pcs_IES_mul_s)

# Incongruent
# cor_pcs_IES_mul_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_MullerLyer_Incongruent)
# 
# IES_pcs_mul_inc_post <- describe_posterior(cor_pcs_IES_mul_inc)
```

#### Vertical Horizontal
```{r}

cor_pcs_IES_vh_con <- BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_VerticalHorizontal_Congruent)

IES_pcs_mul_con_post <- describe_posterior(cor_pcs_IES_vh_con)

cor_pcs_IES_vh_m <- BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_VerticalHorizontal_Mild)

IES_pcs_mul_mild_post <- describe_posterior(cor_pcs_IES_vh_m)

cor_pcs_IES_vh_s <-  BayesFactor::correlationBF(
  y = data_ppt$pcs, 
  x = data_vs_wide$IES_VerticalHorizontal_Strong)

IES_pcs_vh_str_post <- describe_posterior(cor_pcs_IES_vh_s)

# Incongruent
# cor_pcs_IES_vh_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$ErrorRate_VerticalHorizontal_Incongruent)
# 
# IES_pcs_vh_inc_post <- describe_posterior(cor_pcs_IES_vh_inc)
```

## Correlations between Illusion scores and Psychoticism

#### Ebbinghaus 
```{r}

cor_pcs_psy_ebb_con <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_Ebbinghaus_Congruent)

IES_psy_ebb_con_post <- describe_posterior(cor_pcs_psy_ebb_con)

cor_pcs_psy_ebb_m <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_Ebbinghaus_Mild)

IES_psy_ebb_mild_post <- describe_posterior(cor_pcs_psy_ebb_m)

cor_pcs_psy_ebb_s <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_Ebbinghaus_Strong)

IES_psy_ebb_str_post <- describe_posterior(cor_pcs_psy_ebb_s)

# Incongruent
# cor_pcs_psy_ebb_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$IES_Ebbinghaus_Incongruent)
# 
# IES_psy_ebb_inc_post <- describe_posterior(cor_pcs_psy_ebb_inc)
```

#### Muller-Lyer 
```{r}

cor_pcs_psy_mul_con <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_MullerLyer_Congruent)

IES_psy_mul_con_post <- describe_posterior(cor_pcs_psy_mul_con)

cor_pcs_psy_mul_m <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_MullerLyer_Mild)

IES_psy_mul_mild_post <- describe_posterior(cor_pcs_psy_mul_m)

cor_pcs_psy_mul_s <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_MullerLyer_Strong)

IES_psy_mul_str_post <- describe_posterior(cor_pcs_psy_mul_s)

# Incongruent
# cor_pcs_psy_mul_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$IES_MullerLyer_Incongruent)
# 
# IES_psy_mul_inc_post <- describe_posterior(cor_pcs_psy_mul_inc)

#### Vertical-Horizontal
```{r}

cor_pcs_psy_vh_con <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_VerticalHorizontal_Congruent)

IES_psy_vh_con_post <- describe_posterior(cor_pcs_psy_vh_con)

cor_pcs_psy_vh_m <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_VerticalHorizontal_Mild)

IES_psy_vh_mild_post <- describe_posterior(cor_pcs_psy_vh_m)

cor_pcs_psy_vh_s <- BayesFactor::correlationBF(
  y = data_ppt$PID5_Psychoticism,
  x = data_vs_wide$IES_VerticalHorizontal_Strong)

IES_psy_vh_str_post <- describe_posterior(cor_pcs_psy_vh_s)

# Incongruent
# cor_pcs_psy_vh_inc <-  BayesFactor::correlationBF(
#   y = data_ppt$pcs,
#   x = data_vs_wide$IES_VerticalHorizontal_Incongruent)
# 
# IES_psy_vh_inc_post <- describe_posterior(cor_pcs_psy_vh_inc)

>>>>

---
title: "Illusion Sensitivity and Phenomenological Control"
subtitle: "Analysis"
date: "`r Sys.Date()`"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
    warning: false 
    message: false 
execute: 
  cache: false
---

@@

data_ig <- read.csv("../data/data_illusion.csv") |>
  mutate(
    Block = as.factor(Block),
    Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
    Response_Correct = as.factor(Response_Correct)
  )

@@

## Score Computation

### PCS

Mean subjective scale responses for item 4 (Taste) and item 10 (Post-session suggestion) are each generated from the combination of two responses which require processing prior to generating an overall subjective score. 
The subjective score for item 4 is the mean of the responses to the (a)‘sour’ and (b)‘sweet’ sub-scales for the taste suggestion. 
The subjective score for item 10 is the geometric mean (the square root of the product) of he responses to the (a) ‘urge’ and (b) ‘amnesia’ sub-scales for the post-session suggestion.

```{r}
pcs_items <- c("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia")

data_ppt$pcs_tastescore <- rowMeans(data_ppt[grepl("taste", names(data_ppt))])
data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
```

### Illusion Game

```{r}
ig_scores <- data_ig |>
  summarize(ErrorRate = sum(Error) / n(),
            IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
            .by = c("Participant", "Illusion_Type", "Condition_Illusion", "Condition_Difficulty")) |> 
  pivot_wider(names_from = c("Illusion_Type", "Condition_Illusion", "Condition_Difficulty"), 
              values_from = c("ErrorRate", "IES")) 
```

```{r}
corrs <- data.frame()
for(outcome in c("ErrorRate", "IES")) {
  for(ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
    for(cond in c("Congruent", "Mild", "Strong")) {
      corrs <- correlation::correlation(
        select(ig_scores, select=paste0(outcome, "_", ill_type, "_", cond, "_Easy")),
        select(ig_scores, select=paste0(outcome, "_", ill_type, "_", cond, "_Hard"))) |> 
        mutate(Outcome = outcome) |> 
        rbind(corrs)
    }
  }
}

# IF correlations > 0.5:
ig_scores <- data_ig |>
  summarize(ErrorRate = sum(Error) / n(),
            IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
            .by = c("Participant", "Illusion_Type", "Condition_Illusion")) |> 
  pivot_wider(names_from = c("Illusion_Type", "Condition_Illusion"), 
              values_from = c("ErrorRate", "IES")) 
```

#### Manipulation Check

Significant differences between the Congruent and Mild and the Mild and Strong groups are taken as evidence that illusions functioned as intended.

```{r}
ttests <- data.frame()
for(outcome in c("ErrorRate", "IES")) {
  for(ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
    for(cond in c("Congruent", "Mild")) {
      # Add loop for Hard and Easy if need be
      x1 <- ig_scores[[paste0(outcome, "_", ill_type, "_", cond)]]
      if(cond == "Congruent") {
        x2 <- ig_scores[[paste0(outcome, "_", ill_type, "_", "Mild")]]
      } else {
        x2 <- ig_scores[[paste0(outcome, "_", ill_type, "_", "Strong")]]
      }
      ttests <- BayesFactor::ttestBF(x = x1, y = x2, paired = TRUE) |> 
        parameters::parameters() |> 
        mutate(
          Outcome = outcome,
          Illusion_Type = ill_type,
          Condition_Illusion = cond
        ) |> 
        rbind(ttests)
    }
  }
}

ttests |> 
  select(Illusion_Type, Condition_Illusion, Outcome, Median, BF) |> 
  arrange(Illusion_Type, Condition_Illusion, Outcome)
```

## Reliability

### PCS

```{r}
data_ppt[, pcs_items] |>
  psych::alpha()

data_ppt$PCS_Score <- rowMeans(data_ppt[, pcs_items], na.rm = TRUE)
```

### Illusion Game

::: {.panel-tabset}

#### Error Rate - Mild

```{r}
ig_scores |>
  select(matches("ErrorRate_.*_Mild.*")) |>
  psych::alpha()
```

#### Error Rate - Strong

```{r}
ig_scores |>
  select(matches("ErrorRate_.*_Strong.*")) |>
  psych::alpha()
```

#### IES - Mild

```{r}
ig_scores |>
  select(matches("IES_.*_Mild.*")) |>
  psych::alpha()
```

#### IES - Strong

```{r}
ig_scores |>
  select(matches("IES_.*_Strong.*")) |>
  psych::alpha()
```

:::

## Illusion Sensitivity vs. PCS

```{r}
features <- merge(ig_scores, data_ppt[c("Participant", "PCS_Score", "PID5_Psychoticism")])

r <- correlation::correlation(
  select(features, -PCS_Score, -PID5_Psychoticism),
  select(features, PCS_Score),
  bayesian = TRUE
)

r |> 
  format_table() |> 
  select(-Prior, -Method, -n_Obs, -pd, -`% in ROPE`) |> 
  gt::gt() 
```

## Illusion Sensitivity vs. Psychoticism

```{r}
r <- correlation::correlation(
  select(features, -PCS_Score, -PID5_Psychoticism),
  select(features, PID5_Psychoticism),
  bayesian = TRUE
)

r |> 
  format_table() |> 
  select(-Prior, -Method, -n_Obs, -pd, -`% in ROPE`) |> 
  gt::gt()
```
